Adaptive Video Streaming with Optimized Bitstream Extraction and PID-based Quality Control

### 2015-07-25 tmm.tex
rewrite the author line
add more citations to TMM papers

% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
\author{Shengbin~Meng,~Jun~Sun,~Yizhou~Duan~and~Zongming~Guo,~\em{Member,~IEEE} % <-this % stops a space
\thanks{Manuscript received July 15, 2015.}
\thanks{The authors are with the Institute of Computer Science and Technology, Peking University, Beijing%
, 100080, China (Email: shengbin@pku.edu.cn, sunjun@pku.edu.cn, duanyizhou@pku.edu.cn, guozongming@pku.edu.cn; Phone: +86 1082529615; Fax: +86 1082529207). Jun Sun is the corresponding author.}}

@@

\section{Introduction}
\label{sec:intro}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
\IEEEPARstart{V}{ideo} streaming applications are becoming more and more popular on the Internet. Examples of the most common usage scenarios include VOD (Video On Demand), live broadcasting, video-based online education, etc. In these applications, video data is constantly delivered to different clients over various transmission channels, and the client media player can begin playing the video before the entire file has been transmitted. Among other features, video streaming makes it easy to watch video anywhere and anytime, and is an increasingly important way for people to consume video content.

One of the biggest challenge for video streaming is the variety of users' network environment. A user might use different types of network (Ethernet, Wi-Fi, Cellular, etc.) and might also switch from one type to another (e.g., from Wi-Fi to cellular network) on going. Even in the same network, the user's bandwidth will change from time to time, influenced by the overall traffic. Within such an environment, bandwidth fluctuation is inevitable. If the video is streamed at a constant bitrate, it would be difficult or impossible for the user to get best watching experience. To cope with this challenge, the video streaming system should be able to automatically adjust the video's bitrate or quality according to the bandwidth change. In other words, the video streaming system needs to be adaptive.

@@

The advantage of simulcast systems is that they need little modification of existing infrastructure and therefore are easy to deploy. Most commercial video websites, YouTube \cite{YouTube} for instance, use this option to enable quality adaptation. However, the adaptive range and granularity that simulcast provides is limited to the number of streams (YouTube videos have 5 quality levels at most: 240p, 360p, 480p, 720p and 1080p), and the processing and storage of multiple video streams significantly increase time and space cost. SVC streaming systems, on the contrary, do not have above shortcomings. Since an SVC video bitstream can be extracted at the granularity of one NALU (a very small partition of data), it is able to provide seamless scalability and quality adaptation. Besides, performance analysis \cite{SVCPerformance} shows that SVC also has much higher coding efficiency than simulcast, despite some computation and bitrate overhead comparing with single layer stream. Therefore, to adopt SVC in the video streaming framework is a promising option and has attracted much attention \cite{Zhu2013}\cite{Dan2013}\cite{Yang2014}.

In the SVC adaptive video streaming system, there are two key problems that need to be addressed well. First, as the prerequisite of the adaption to bandwidth change, the whole bitstream needs to be extracted so that sub-streams with various bitrates can be obtained. This is called the bitstream extraction problem. Optimal bitstream extraction algorithm aims to extract a sub-stream with best video quality, or minimal distortion, under the constraint of a given bitrate. Second, when the bandwidth changes with time, it's necessary to make decisions about whether and how to adjust the bitrate or video quality. This problem is called quality control, which means controlling the video quality (or the video bitrate, from the transmission point of view) to follow the change of the bandwidth. A good quality control algorithm should determine a suitable bitrate for current bandwidth condition, and keep the video quality both smooth and at a high level.

@@

In (\ref{eq:ut}), $K_p$, $K_i$, and $K_d$ indicate the tuning parameters for proportional, integral and derivative parts respectively, and $E_p^t$, $E_i^t$, $E_d^t$ stand for the corresponding observed value for each part, which are defined in (\ref{eq:ep}) - (\ref{eq:long-check}); the controller output $u_t$ is utilized to calculate an appropriate data rate and further to determine which quality level we should transmit (see Algorithm \ref{algo:control} for more details).

@@

@@

\bibitem{SVCOverview}
H.~Schwarz, D.~Marpe and T.~Wiegand, ``Overview of the scalable video coding extension of the H.264/AVC standard'', in {\em IEEE Trans. Circuits. Syst. Video Technol.}, vol. 17, no. 9, pp.~1103-1120, Sept. 2007.

@@

\bibitem{Zhu2013}
Z. Zhu, S. Li, and X. Chen, ``Design QoS-aware multi-path provisioning strategies for efficient cloud-assisted SVC video streaming to heterogeneous clients'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 4, pp.758-768, 2013.

\bibitem{Dan2013}
G. Dan, O. Hadar, R. Ohayon, and N. Amram, ``Live video streaming with adaptive pre-processing by using scalable video coding'', in {\em Proceedings of IEEE International Conference on Consumer Electronics (ICCE)}, pp.588-589, 2013.

\bibitem{Yang2014}
E. Yang, Y. Ran, S. Chen, and J. Yang, ``A multicast architecture of SVC streaming over OpenFlow networks'', in {\em Proceedings of IEEE Global Communications Conference (GLOBECOM)}, pp.1323-1328, 2014.

@@

%\bibitem{Meng14}
%Shengbin Meng, Jun Sun, Yizhou Duan, Zongming Guo, ``An Efficient Method For No-Reference H.264/SVC Bitstream Extraction'', {\em IEEE International Conference on Acoustics, Speech, and Signal Processing}, Florence, Italy, May, 2014.

%\bibitem{Meng14-PID}
%Shengbin Meng, Jun Sun, Yilei Wang, Zongming Guo, ``A PID-based Quality Control Algorithm for SVC Video Streaming'', {\em IEEE International Conference on Communications}, Sydney, Australia, June, 2014.

@@

\bibitem{HEVC}
G.~J.~Sullivan, J.-R.~Ohm, W.-J.~Han, and T.~Wiegand, ``Overview of the High Efficiency Video Coding (HEVC) standard'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol.~22, no.~12, pp.~1648–1667, Dec. 2012.

>>>>

% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
\author{Shengbin~Meng, Jun~Sun, {\em Member, IEEE}, Yizhou~Duan and Zongming~Guo, {\em Member, IEEE} % <-this % stops a space
\thanks{Manuscript received July 15, 2015.}
\thanks{The authors are with the Institute of Computer Science and Technology, Peking University, Beijing%
, 100080, China (Email: shengbin@pku.edu.cn, sunjun@pku.edu.cn, duanyizhou@pku.edu.cn, guozongming@pku.edu.cn; Phone: +86 1082529615; Fax: +86 1082529207). Jun Sun is the corresponding author.}}

@@

\section{Introduction}
\label{sec:intro}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
\IEEEPARstart{V}{ideo} streaming applications are becoming more and more popular on the Internet. Examples of the most common usage scenarios include online video-on-demand (VoD), live broadcasting, video chat, video-based online education, etc. In these applications, video data is constantly delivered to different clients over various transmission channels, and the client media player can begin playing the video before the entire file has been transmitted. Among other features, video streaming makes it easy to watch video anywhere and anytime, which represents the concept of Ubiquitous Multimedia Access (UMA), a very common topic within the multimedia community during the past few years \cite{Gualdi08}. And due to its convenience, video streaming has also attracted a large number of users and become an increasingly important way for people to consume video content \cite{Chen13}.

Although efforts on the new software defined networking (SDN) are made to facilitate efficient video streaming \cite{Egilmez13}, currently traditional best-effort networks still dominate, and the variety of users' network environment remains to be one of the biggest challenges for video streaming. A user might use different types of network (Ethernet, Wi-Fi, Cellular, etc.) and might also switch from one type to another (e.g., from Wi-Fi to cellular network) on going. Even in the same network, the user's bandwidth will change from time to time, influenced by the overall traffic. Within such an environment, bandwidth fluctuation is inevitable. If the video is streamed at a constant bitrate, it would be difficult or impossible for the user to get best watching experience. To cope with this challenge, the video streaming system should be able to automatically adjust the video's bitrate or quality according to the bandwidth change. In other words, the video streaming system needs to be adaptive.

@@

The advantage of simulcast systems like DASH is that they need little modification of existing infrastructure (servers, proxies, caches, firewall compatibility, etc.) and therefore are easy to deploy \cite{Bouten14}. Most commercial video websites, YouTube \cite{YouTube} for instance, use this option to enable quality adaptation. However, the adaptive range and granularity that simulcast provides is limited to the number of streams (YouTube videos have 5 quality levels at most: 240p, 360p, 480p, 720p and 1080p), and the processing and storage of multiple video streams significantly increase time and space cost. SVC streaming systems, on the contrary, do not have above shortcomings. Since an SVC video bitstream can be extracted at the granularity of one NALU (a very small partition of data), it is able to provide seamless scalability and quality adaptation. Besides, performance analysis \cite{SVCPerformance} shows that SVC also has much higher coding efficiency than simulcast, despite some computation and bitrate overhead comparing with single layer stream. Therefore, to adopt SVC in the video streaming framework is a promising option and has attracted much attention in the study \cite{Chuah12}-\cite{Cicalò14}.

While these studies mostly focus on Quality of Service (QoS) optimization and resource allocation for specific networks, there are two key problems in the SVC adaptive video streaming system that need to be primarily addressed. First, as the prerequisite of the adaption to bandwidth change, the whole bitstream needs to be extracted so that sub-streams with various bitrates can be obtained. This is called the bitstream extraction problem. Optimal bitstream extraction algorithm aims to extract a sub-stream with best video quality, or minimal distortion, under the constraint of a given bitrate. Second, when the bandwidth changes with time, it's necessary to make decisions about whether and how to adjust the bitrate or video quality. This problem is called quality control, which means controlling the video quality (or the video bitrate, from the transmission point of view) to follow the change of the bandwidth. A good quality control algorithm should determine a suitable bitrate for current bandwidth condition, and keep the video quality both smooth and at a high level.

@@

In (\ref{eq:ut}), $K_p$, $K_i$, and $K_d$ indicate the tuning parameters for proportional, integral and derivative parts respectively, and $E_p^t$, $E_i^t$, $E_d^t$ stand for the corresponding observed value for each part, which are defined in (\ref{eq:ep})-(\ref{eq:long-check}); the controller output $u_t$ is utilized to calculate an appropriate data rate and further to determine which quality level we should transmit (see Algorithm \ref{algo:control} for more details).

@@

\bibitem{Gualdi08}
G. Gualdi, A. Prati, and R. Cucchiara, ``Video Streaming for Mobile Video Surveillance'', in {\em IEEE Transactions on Multimedia}, vol. 10, no. 6, pp.1142-1154, 2008.

\bibitem{Chen13}
Y. Chen, B. Zhang, Y. Liu, and W. Zhu, ``Measurement and Modeling of Video Watching Time in a Large-Scale Internet Video-on-Demand System'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 8, pp.2087-2098, 2013.

\bibitem{Egilmez13}
H. E. Egilmez, S. Civanlar, and A. M. Tekalp, ``An Optimization Framework for QoS-Enabled Adaptive Video Streaming Over OpenFlow Networks'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 3, pp.710-715, 2013.

@@

\bibitem{SVCOverview}
H.~Schwarz, D.~Marpe, and T.~Wiegand, ``Overview of the scalable video coding extension of the H.264/AVC standard'', in {\em IEEE Trans. Circuits. Syst. Video Technol.}, vol. 17, no. 9, pp.~1103-1120, Sept. 2007.

\bibitem{Bouten14}
N. Bouten, S. Latré, J. Famaey, W. V. Leekwijck, and F. D. Turck, ``In-Network Quality Optimization for Adaptive Video Streaming Services'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 8, pp.2281-2293, 2014.

@@

\bibitem{Chuah12}
S. Chuah, Z. Chen, and Y. Tan, ``Energy-Efficient Resource Allocation and Scheduling for Multicast of Scalable Video Over Wireless Networks'', in {\em IEEE Transactions on Multimedia}, vol. 14, no. 4, pp.1324-1336, 2012.

\bibitem{Zhu13}
Z. Zhu, S. Li, and X. Chen, ``Design QoS-aware multi-path provisioning strategies for efficient cloud-assisted SVC video streaming to heterogeneous clients'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 4, pp.758-768, 2013.

\bibitem{Dan13}
G. Dan, O. Hadar, R. Ohayon, and N. Amram, ``Live video streaming with adaptive pre-processing by using scalable video coding'', in {\em Proceedings of IEEE International Conference on Consumer Electronics (ICCE)}, pp.588-589, 2013.

\bibitem{Yang14}
E. Yang, Y. Ran, S. Chen, and J. Yang, ``A multicast architecture of SVC streaming over OpenFlow networks'', in {\em Proceedings of IEEE Global Communications Conference (GLOBECOM)}, pp.1323-1328, 2014.

\bibitem{Cicalò14}
S. Cicalò, and V. Tralli, ``Distortion-Fair Cross-Layer Resource Allocation for Scalable Video Transmission in OFDMA Wireless Networks'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 3, pp.848-863, 2014.

@@

@@

\bibitem{HEVC}
G.~J.~Sullivan, J.-R.~Ohm, W.-J.~Han, and T.~Wiegand, ``Overview of the High Efficiency Video Coding (HEVC) standard'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol.~22, no.~12, pp.~1648â1667, Dec. 2012.
